---
title: "mGPS - MetaSUB global AMR analysis and figures"
author: "Yali Zhang and Eran Elhaik"
output: github_document
---

```{r setup, echo = F}
### path to mGPS top-level directory here ###
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width=13, fig.height=8) 
```

This notebook outlines the procedure used to generate MetaSUB global AMR analysis for the mGPS paper, including the code used for analysis and figures.  

Load dependencies

```{r, echo=T, eval = T, message = F}
library(ggplot2)
library(ggsignif)
library(forcats)
library(ggpmisc)
```


```{r, echo = F}
metasub_data <- read.csv("Metasub/Outputs/Global/metasub_global_results.csv")
AMR_data <- read.csv('Data/Metasub/After_process_megares_amr_class_rpkmg.csv', check.names = F)

metasub_ARM<- merge(metasub_data,AMR_data,by.x="uuid",by.y="uuid")
```

Calculate relative sequence abundance of AMR for local and non-local samples

```{r}
AMR_list <- colnames(AMR_data)[2:22]

metasub_ARM$city <- factor(metasub_ARM$city)

bar_df_abundance <- data.frame(row.names = c(levels(metasub_ARM$city),paste0(levels(metasub_ARM$city),':migrate')))

for (i in 1:length(levels(metasub_ARM$city))){
  
  for (j in 1: length(AMR_list)){
    
    city <- levels(metasub_ARM$city)[i]
    taxa <- AMR_list[j]  
    city_subset <- subset(metasub_ARM[metasub_ARM$city == city,])
    origin_same <- subset(city_subset[city_subset$city == city_subset$cityPred,])
    origin_diff <- subset(city_subset[city_subset$city != city_subset$cityPred,])
    
    abun_same <- sum(origin_same[,taxa])/length(origin_same[,taxa])
    abun_diff <- sum(origin_diff[,taxa])/length(origin_diff[,taxa])
    
    bar_df_abundance[city,taxa] <- abun_same
    
    migrate <- paste0(city,':migrate')
    bar_df_abundance[migrate,taxa] <- abun_diff
  }
}

# On glocal level
global_native <- c()
global_non <- c()
global_city_native <- c()
global_city_non <- c()
for (j in 1:length(AMR_list)){
  taxa <- AMR_list[j]  
  
  origin_same <- subset(metasub_ARM[metasub_ARM$city == metasub_ARM$cityPred,])
  origin_diff <- subset(metasub_ARM[metasub_ARM$city != metasub_ARM$cityPred,])
  
  abun_same <- mean(origin_same[,taxa])
  abun_diff <- mean(origin_diff[,taxa])
  
  global_native <- c(global_native,abun_same)
  global_non <- c(global_non,abun_diff)
  
  global_city_native <- c(global_city_native, mean(bar_df_abundance[c(1:38,40),taxa]))
  global_city_non <- c(global_city_non,mean(bar_df_abundance[41:80,taxa]))
}

```

Fig 8-A Comparing RPKM values of each AMRs between local and non-local samples on a global level. Bar colors distinguish local (green) from non-local (red) samples.

```{r}
global_avg_df_1 <- cbind.data.frame('AMR'= AMR_list,'Source'=rep('Local',21),'value'=global_native)
global_avg_df_2 <- cbind.data.frame('AMR'= AMR_list,'Source'=rep('Non-local',21),'value'=global_non)
global_avg_df <- rbind.data.frame(global_avg_df_1,global_avg_df_2)

global_avg_df$AMR <- factor(global_avg_df$AMR)
global_avg_df$AMR <- fct_inorder(global_avg_df$AMR)
global_avg_df$AMR <- fct_relevel(global_avg_df$AMR,
                                         rev(levels(global_avg_df$AMR))) 

global_avg_df$Source <- factor(global_avg_df$Source)
global_avg_df$Source <- fct_relevel(global_avg_df$Source,
                                 rev(levels(global_avg_df$Source))) 

ggplot(global_avg_df,aes(x=AMR,y=value,fill=Source))+
  geom_bar(stat="identity",position=position_dodge(0.75))+
  coord_flip()+
  ylab('AMR relative abundance (RPKM)')+
  theme_bw()+
  theme(title=element_text(size=8),
        axis.text=element_text(size=8),
        legend.text = element_text(size =8))

```

Fig 8-B Comparisons of AMRs’ RPKM values distributions of local and non-local samples globally.

```{r}

#global individual
for (i in 1:length(metasub_ARM$uuid)){
 if (metasub_ARM[i,'city'] == metasub_ARM[i,'cityPred']){
   metasub_ARM[i,'Migration'] = 'Local'
 }else{
   metasub_ARM[i,'Migration'] = 'Non-local'
 }
}

AMR_df <- subset(metasub_ARM[,c(AMR_list[1:20],'Migration')])
reshape2::melt(AMR_df,id.vars="Migration") -> dfa0

dfa0$Migration <- factor(dfa0$Migration)

levels(dfa0$Migration) <- c('Local','Non-local')

ggplot(dfa0, aes(x=Migration,y=value,fill=Migration)) + 
  geom_boxplot() +
  ylim(0,100)+
  ylab('AMR relative abundance (RPKM)')+
  xlab('Sample source')+
  guides(fill=FALSE) +
  # ggtitle('AMR comparison for local and non-local samples (wilcox.test)')+
  facet_wrap(~variable) +
  geom_signif(comparisons = list(c("Local", "Non-local")),
              step_increase = 0.3,
              map_signif_level = T,
              test = wilcox.test)+
  theme_bw()
```

Fig 8-C Total AMR RPKM values for local and non-local samples per city. 

```{R}
bar_df_abundance2 <- bar_df_abundance
bar_df_abundance2$city_1 <- row.names(bar_df_abundance)
reshape2::melt(bar_df_abundance2,id.vars="city_1") -> dfa1

for (i in 1:length(dfa1$city_1)){
  city <- dfa1$city_1[i]
  
  if (':' %in% strsplit(city,'')[[1]]){
    country <- strsplit(city,':')[[1]][1]
    
    if ('_' %in% strsplit(country,'')[[1]]){
      
      country <- strsplit(country,'_')[[1]]
      
      for (a in 1:length(country)){
        country_a <- strsplit(country[a],'')[[1]]
        country_a[1] <- toupper(country_a[1])
        country_a <- paste(country_a, collapse='')
        country[a] <- country_a}
      country <- paste(country, collapse=' ')
      
    }else{
      
      country <- strsplit(country,'')[[1]]
      country[1] <- toupper(country[1])
      country <- paste(country, collapse='')
    }
    
    dfa1[i,'city'] <- country
    dfa1[i,'Sample origin'] <- 'Non-native'
  }else{
    
    country <- city
    
    if ('_' %in% strsplit(country,'')[[1]]){
      
      country <- strsplit(country,'_')[[1]]
      
      for (a in 1:length(country)){
        country_a <- strsplit(country[a],'')[[1]]
        country_a[1] <- toupper(country_a[1])
        country_a <- paste(country_a, collapse='')
        country[a] <- country_a}
      country <- paste(country, collapse=' ')
      
    }else{
      
      country <- strsplit(country,'')[[1]]
      country[1] <- toupper(country[1])
      country <- paste(country, collapse='')
    
    dfa1[i,'city'] <- country
    dfa1[i,'Sample origin'] <- 'Native'
  }
  
}

# AMR_Sum analysis for each city
dfa1$variable <- as.character(dfa1$variable)
ARM_sum_df <- subset(dfa1[dfa1$variable == 'AMR (total)',])

city_anno <- read.csv('Data/Metasub/MetaSUB City Metadata - Sheet1 new.csv')

for (i in 1:length(ARM_sum_df$city)){
  city <- ARM_sum_df$city_1[i]
  if (':' %in% strsplit(city,'')[[1]]){
    city <- strsplit(city,':')[[1]][1]
  }
  ARM_sum_df[i,'country'] <- city_anno[city_anno$city == city,'country']
  ARM_sum_df[i,'continent'] <- city_anno[city_anno$city == city,'continent']
}

ARM_sum_df <- ARM_sum_df[order(ARM_sum_df$country),]
ARM_sum_df <- ARM_sum_df[order(ARM_sum_df$continent),]

for (i in 1:length(ARM_sum_df$city_1)){
  if (':' %in% strsplit(ARM_sum_df$city_1[i],'')[[1]]){
    ARM_sum_df[i,'order'] <- ARM_sum_df[i,'value']
    city_m <- strsplit(ARM_sum_df$city_1[i],':')[[1]][1]
    ARM_sum_df[ARM_sum_df$city_1 == city_m,'order'] <- ARM_sum_df[i,'value']
  }
}

ARM_sum_df2 <- ARM_sum_df[order(ARM_sum_df$order),]
ARM_sum_df2$city <- factor(ARM_sum_df2$city)
ARM_sum_df2$city <- fct_inorder(ARM_sum_df2$city)
ARM_sum_df2$city <- fct_relevel(ARM_sum_df2$city,rev(levels(ARM_sum_df2$city))) 
ARM_sum_df3 <- ARM_sum_df2
ARM_sum_df3[length(ARM_sum_df3$city_1),'value'] <- 10

ggplot(ARM_sum_df3,aes(x=city,y=value,fill=`Sample origin`))+
  geom_bar(stat="identity",position=position_dodge(1))+
  ylab('AMR relative abundance (RPKM)')+
  xlab('City')+
  theme_bw()+
  theme(axis.text.x = element_text( vjust = 0.5, hjust = 1, angle = 90))+
  annotate("text", x=4.9 , y=10,label = "* True value = 13.65", )+
  scale_fill_manual(values = c("#00BFC4", "#F8766D"),
                    breaks = c("Native", "Non-native"),
                    label = c('Local','Non-local'),
                    name = 'Source')+
  theme(legend.background = element_rect(
    size=0.3, linetype="solid",
    colour ="black"))

```

Fig 8-D The correlation of total AMR RPKM values of non-local samples and the annual number of tourists

```{r}
tour_df <- ARM_sum_df[ARM_sum_df$`Sample origin`=='Non-native',]

for (i in 1:length(tour_df$city_1)){
  city <- strsplit(tour_df$city_1[i],':')[[1]][1]
  tour_df[i,'tourists'] <- city_anno[city_anno$city == city,'number_of_tourists_annually']
}

tour_df2 <- tour_df[order(tour_df$value,decreasing = T),]
tour_df2 <- tour_df2[-1,]

options(scipen=200)

ggplot(data = tour_df,aes(x=tourists,y=value))+
  geom_point(size=2)+
  ylab('AMR relative abundance (RPKM)')+
  xlab('Number of tourists annually')+
  stat_smooth(method=lm,formula=y~x,se=FALSE)+
  geom_smooth (method = lm,linetype=1,se=FALSE,span=1)+
  theme_bw()+
  theme(axis.text.x = element_text( vjust = 0.5, hjust = 1))
```
